flowers-pilot-analysis

### 2021-07-07 code/aa-flowers-pilot_analysis.Rmd
add technical pilots

```{r set-up, include=F}
#knitr::opts_chunk$set(echo = FALSE, warning=F, message=F)
#knitr::opts_chunk$set(dev = "png", dev.args = list(type = "cairo-png"))
#options(knitr.table.format = "html")
library(tidyverse)
library(jsonlite)
library(here)
#library(rlang)
#library(lme4)
#library(brms)
#library(rstan)
#rstan_options(auto_write = TRUE)
#options(mc.cores = parallel::detectCores())
theme_set(theme_bw())
ParseJSONColumn <- function(x) {
  str_c("[ ", str_c(x, collapse = ",", sep=" "), " ]")  %>% 
    fromJSON(flatten = T)
}
data_target = "pilot0"
##Data import constants
data_location=paste0("data/",data_target)
date_start=lubridate::ymd('2021-05-27')
image_location=paste0("figs/",data_target)
#model_location="code/models"
```

@@

```{r}
extract_context_info <- function(context_key){
  return(str_extract(context_key, "[^.]*$"))
}

d.contexts <- read_csv(here(data_location,'rounds.csv'),guess_max=10000) %>% 
  mutate(data.context = map(data.context, 
                            .f = ParseJSONColumn)) %>% 
  select("_id":"data.numPlayers") %>%
  mutate(data.context = map(data.context,
                            function(df) df %>% gather(key, value) %>%
                              mutate(key = gsub('.', '-', key, fixed=TRUE)) %>%
                              separate(key, into = c("label", "key"), sep = "-") %>%
                              spread(key, value))) %>%
  unnest(data.context)

d.games <- read_csv(here(data_location, 'games.csv')) %>% 
  rename(gameId = `_id`) %>% 
    filter(createdAt >= date_start)

d.chat.raw <- read_csv(here(data_location, 'rounds.csv'), guess_max=10000) %>%
  filter(createdAt >= date_start) %>%
  mutate(data.chat = ifelse(is.na(data.chat), '{}', data.chat)) %>%
  rename(row_id = `_id`) %>%
  mutate(data.chat = map(data.chat, .f = ParseJSONColumn)) %>%
  unnest(data.chat) %>%
  select(-ends_with('response'), -ends_with('time')) %>%
  rename_with(~ gsub("data.", "", .x, fixed = TRUE)) %>% 
  write_csv(here(data_location, 'raw_chat.csv'))

d.round_results.raw <- read_csv(here(data_location,'rounds.csv'),guess_max=10000) %>% 
  filter(createdAt >= date_start) %>% 
  rename_with(~ gsub("data.", "", .x, fixed = TRUE)) %>% 
  rename_with ( ~ gsub("room", "player", .x, fixed=T)) %>% 
  select(-chat) %>% 
  gather(key, value, starts_with('player')) %>% 
  separate(key, into = c('blah', 'playerId', 'info')) %>% 
  spread(info, value) %>% 
  select(-blah, -context) %>% 
  left_join(d.contexts %>% select(gameId, 
                                  repNum = data.repNum, 
                                  blockNum = data.blockNum, 
                                  trialNum = data.trialNum, 
                                  response = label, utility, blinded)) %>%
  mutate(time = as.numeric(time)/1000) %>%
  #get individual utilities 
  group_by(gameId, repNum, blockNum, trialNum, response) %>% 
  mutate(n_resp = n(), ind_winnings = as.numeric(utility)*n_resp)

#only include rounds that finished
rounds_exclude <- d.round_results.raw %>% group_by(gameId,numPlayers,repNum) %>% tally() %>% filter(n!=12) %>% select(gameId,repNum)
d.round_results <-  d.round_results.raw %>% anti_join(rounds_exclude)
```

```{r}
summary <- d.round_results.raw %>% group_by(trialNum, repNum, gameId, numPlayers) %>% 
          # mutate(time= time %|% 180) %>% 
  summarize(max_time=max(time)) %>% 
  group_by(gameId, numPlayers) %>% 
  summarize(total_time=sum(max_time)/60,
            num_rounds=max(repNum+1)) %>% 
  arrange(numPlayers)
#knitr::kable(summary)
message("Full games")
summary %>% filter(num_rounds==12) %>% group_by(numPlayers) %>% tally()
message("Partial games")
summary %>% filter(num_rounds!=12) %>% group_by(numPlayers) %>% tally()
```

@@

```{r chat}
#group raw chat by participant
d.chat <- read_csv(here(data_location, "raw_chat.csv")) %>% 
 # filter(!is.chitchat) %>% 
  mutate(text = gsub("\\n", '', fixed = T, text),
         text = gsub("[/?/.]", ' ', text),
         text = str_squish(text),
         #tangram = gsub('/experiment/tangram_', '', target, fixed=TRUE),
         #tangram = gsub('.png', '', tangram, fixed=TRUE),
         utt_length_chars = str_length(text), 
         utt_length_words = str_count(text, "\\W+") + 1) %>% 
  filter(type != "alert")%>%
  group_by(gameId, blockNum, trialNum, repNum, playerId, numPlayers) %>%
  summarize(text = paste0(text, collapse = ', '),
            total_num_words = sum(utt_length_words),
            total_num_chars = sum(utt_length_chars)) %>%
  anti_join(rounds_exclude) %>% 
  full_join(d.round_results, c("gameId", "trialNum", "repNum", "playerId", "numPlayers", "blockNum"))
```

>>>>

```{r set-up, include=F}
#knitr::opts_chunk$set(echo = FALSE, warning=F, message=F)
#knitr::opts_chunk$set(dev = "png", dev.args = list(type = "cairo-png"))
#options(knitr.table.format = "html")
library(tidyverse)
library(jsonlite)
library(here)

theme_set(theme_bw())

data_target = "technical_pilots"
##Data import constants
data_location=paste0("data/",data_target)
date_start=lubridate::ymd('2021-06-30')
image_location=paste0("figs/",data_target)

#library(rlang)
#library(lme4)
#library(brms)
#library(rstan)
#rstan_options(auto_write = TRUE)
#options(mc.cores = parallel::detectCores())
#model_location="code/models"

@@

We need:
1) game info with treatment parameters
2) context for each round (what were participant's options? max utility based on game conditions?)
3) round results (what did participants select each round?)
4) chat (what did individual participants say each round)
5) feedback/demographics

```{r helper-functions}
ParseJSONColumn <- function(x) {
  str_c("[ ", str_c(x, collapse = ",", sep=" "), " ]")  %>% 
    fromJSON(flatten = T)
}

extract_context_info <- function(context_key){
  return(str_extract(context_key, "[^.]*$"))
}
```

Get info about the games and treatments
- We need each game, the 
```{r game-preprocessing}
d.games.raw <- read_csv(here(data_location, 'games.csv')) %>% 
  rename(gameId = `_id`) %>% 
  filter(createdAt >= date_start) %>%
  rename_with(~ gsub("data.", "", .x, fixed = TRUE))

d.treatments.raw <- read_csv(here(data_location, 'treatments.csv')) %>% 
  rename(treatmentId = `_id`) %>%
  #filter(createdAt >= date_start) %>%
  mutate(factorIds = str_split(factorIds, pattern=",")) %>%
  unnest(factorIds) %>% rename(factorId = factorIds)

d.factors.raw <- read_csv(here(data_location, 'factors.csv'))%>% 
  rename(factorId = `_id`)

d.factor_types.raw <- read_csv(here(data_location, 'factor-types.csv')) %>%
  rename(factorTypeId = "_id",
         factorName = "name")

d.factors <- d.factors.raw %>% 
  left_join(d.factor_types.raw %>% 
              select(factorTypeId, factorName))

d.treatments <- d.treatments.raw %>% 
  left_join(d.factors %>% 
              select(factorId, factorName, value)) %>%
  pivot_wider(id_cols = c(treatmentId, name, createdAt), 
              names_from = factorName, values_from = value)

d.players <- read_csv(here(data_location, 'players.csv')) %>%
  rename(playerId = `_id`) %>% 
  filter(createdAt >= date_start) %>%
  rename_with(~ gsub("data.", "", .x, fixed = TRUE))

d.games <- d.games.raw %>% left_join(d.treatments %>% 
                                       select(-createdAt)) %>%
  janitor::remove_empty() %>%
  #select out the nonuseful columns
  select(-c(gameLobbyId, treatmentId, batchId, targetSet, justStarted, submitCode, roundIds)) %>%
  separate(col = playerIds, sep = ",", into = c("player_1", "player_2", "player_3")) %>% 
  pivot_longer(cols = starts_with("player_"), names_to = "playerNum", values_to = "playerId") %>%
  left_join(d.players %>% select(playerId, readyAt)) %>%
  mutate(game_length = finishedAt - readyAt) %>%
  group_by(across(gameId:chatEnabled)) %>%
  summarize(gameLength = max(game_length)) %>% ungroup()
```

Round info
For each round, we want to know the context...
- which flowers were available
- the utility of each flower

We also want to have a report for each round including...
- which participants selected which flowers
- the time the participants took to select
- the utility of the flower the participants selected

```{r}
d.round_results.raw <- read_csv(here(data_location,'rounds.csv'), guess_max=10000) %>% 
  filter(createdAt >= date_start) %>% 
  rename_with(~ gsub("data.", "", .x, fixed = TRUE)) %>% 
  rename_with ( ~ gsub("room", "player", .x, fixed=T))

#try pivot_longer instead of gather??
d.round_results <- d.round_results.raw  %>% 
  select(-chat, -context) %>%
  mutate_at(vars(starts_with('player')), as.character) %>%
  pivot_longer(cols = starts_with('player'), 
               names_to = c('blah', 'playerId', 'info'), 
               values_to = "value",
               names_sep = "_",
               values_drop_na = TRUE) %>%
  pivot_wider(names_from = "info", values_from = "value") %>%
  select(-blah) %>%
  mutate_at(c("time", "utility"), as.numeric) %>%
  #TODO: some times are negative?? interesting
  mutate(time = abs(as.numeric(time)/1000)) %>%
  rename(time_sec = time, playerResponse = response, playerUtility = utility) %>%
  arrange(createdAt, gameId, blockNum, repNum)

#exclude blocks with <12 completed rounds
rounds_exclude <- d.round_results %>% group_by(gameId, blockNum) %>% tally() %>% filter(n!=36) %>% select(gameId,blockNum)
d.round_results.final <-  d.round_results %>% anti_join(rounds_exclude)
```
```{r}
summary <- d.round_results %>% 
  group_by(gameId) %>% 
  summarize(num_rounds=max(repNum+1)) %>% 
  left_join(d.games%>% ungroup() %>% select(gameId, condition, chat=chatEnabled, gameLength))
#knitr::kable(summary)
message("Full games")
summary %>% filter(num_rounds==12)
message("Partial games")
summary %>% filter(num_rounds!=12) 
```

```{r}
d.chat.raw <- d.round_results.raw %>%
  filter(createdAt >= date_start) %>%
  anti_join(rounds_exclude) %>%
  #there will always be an alert for each active round, so drop NA
  filter(!is.na(chat))%>%
  select(-c(ends_with('response'), ends_with('time'), ends_with('utility'), context, stageIds)) %>%
  #mutate(chat = ifelse(is.na(chat), '{}', chat)) %>%
  rename(roundID = `_id`) %>%
  mutate(chat = map(chat, .f = ParseJSONColumn)) %>%
  unnest(chat) %>%
  arrange(createdAt, gameId, blockNum, repNum) %>% 
  left_join(d.players %>% select(playerId, name)) %>%
  left_join(d.round_results %>% select(gameId, blockNum, repNum, trialNum, playerId, playerResponse, playerUtility)) %>%
  left_join(d.games %>% select(gameId, condition, chatEnabled)) %>%
  mutate("participantAction" = ifelse(type == "alert", paste(name, "selected", playerResponse, "for", playerUtility, "point(s)"), NA)) %>%
  select(gameId, trialNum, condition, chatEnabled, playerId, name, text, participantAction, everything()) %>%
  write_csv(here(data_location, 'raw_chat.csv'))

#extract the alternative flowers
d.contexts <- read_csv(here(data_location,'rounds.csv'),guess_max=10000) %>% 
  mutate(data.context = map(data.context, 
                            .f = ParseJSONColumn)) %>% 
  select("_id":"data.numPlayers") %>%
  mutate(data.context = map(data.context,
                            function(df) df %>% gather(key, value) %>%
                              mutate(key = gsub('.', '-', key, fixed=TRUE)) %>%
                              separate(key, into = c("label", "key"), sep = "-") %>%
                              spread(key, value))) %>%
  unnest(data.context) %>%
  rename_with(~ gsub("data.", "", .x, fixed = TRUE))
```
How often do participants select a flower that is hidden from them?

@@

Correlation between response time and utility?

```{r}
d.round_results %>% left_join(d.games %>% select(gameId, condition, chatEnabled)) %>%
  mutate(chat = ifelse(chatEnabled, "lang", "nonlang"),
         full_condition = paste0(condition, "-", chat)) %>%
  ggplot(aes(x=log(time_sec), y = as.numeric(playerUtility), color = full_condition)) + 
  facet_grid(rows = vars(condition), cols = vars(chat)) +
  geom_point(alpha = .25) + 
  labs(x = "log time (in seconds)",
       y = "individual utility",
       title = "Response time in sec (log) and utility", subtitle =  "each participant, each round")
```

Easiest way to do this, is to derive the game start date with the latest participant ready time and the game finishedAt

```{r}
d.round_results %>% left_join(d.games %>% select(gameId, condition, chatEnabled, gameLength)) %>%
  mutate(chat = ifelse(chatEnabled, "lang", "nonlang"),
         full_condition = paste0(condition, "-", chat)) %>%
  group_by(gameId, condition, chat, full_condition,gameLength) %>%
  summarize(totalUtility = sum(playerUtility))%>%
  ggplot(aes(x=gameLength, y = as.numeric(totalUtility), color = full_condition)) + 
  facet_grid(rows = vars(condition), cols = vars(chat), scales = "free_y") +
  geom_point(alpha = .75) + 
  labs(x = "Game length (in minutes)",
       y = "total points by group",
       title = "Length of game in mins and utility", subtitle =  "per game")

d.round_results %>% left_join(d.games %>% select(gameId, condition, chatEnabled, gameLength)) %>%
  mutate(chat = ifelse(chatEnabled, "lang", "nonlang"),
         full_condition = paste0(condition, "-", chat)) %>%
  group_by(gameId, condition, chat, full_condition,gameLength) %>%
  summarize(totalUtility = sum(playerUtility))%>%
  mutate(totalUtility = ifelse(condition=="coopCartel", totalUtility/3, totalUtility)) %>%
  ggplot(aes(x=log(as.numeric(gameLength, units="mins")), y = as.numeric(totalUtility), color = full_condition)) + 
  #facet_grid(rows = vars(condition), cols = vars(chat), scales = "free_y") +
  geom_point(alpha = .75) + 
  geom_smooth(method=lm, alpha=.3)+
  labs(x = "Game length (in minutes)",
       y = "total points by group",
       title = "Length of game in mins and utility", subtitle =  "per game")
```

```{r chat}
#group raw chat by participant
d.chat <- read_csv(here(data_location, "raw_chat.csv"), guess_max=1000) %>% 
  filter(type == "message") %>%
  full_join(d.round_results, c("gameId", "trialNum", "repNum", "playerId", "numPlayers", "blockNum")) %>%
 # filter(!is.chitchat) %>% 
  mutate(text = gsub("\\n", '', fixed = T, text),
         text = gsub("[/?/.]", ' ', text),
         text = str_squish(text),
         utt_length_chars = str_length(text), 
         utt_length_words = str_count(text, "\\W+") + 1) %>% 
  group_by(gameId, blockNum, trialNum, repNum, playerId, numPlayers) %>%
  summarize(text = paste0(text, collapse = ', '),
            total_num_words = sum(utt_length_words),
            total_num_chars = sum(utt_length_chars)) %>%
  anti_join(rounds_exclude)

Pick just one game to look at, make sure the plot is working?
```{r}
condition = "coopCartel"

d.chat %>% 
         left_join(d.games %>% select(gameId, createdAt,condition, chatEnabled)) %>% 
         filter(condition == "coopCartel",
                chatEnabled == "true") %>%
ggplot(aes(x = repNum, y=total_num_words, color = as.factor(playerId))) + 
  geom_jitter(alpha = .5) + 
  facet_grid(cols = vars(blockNum), rows = vars(gameId), scales = "free_y") + 
  geom_smooth(method=glm, formula=y~poly(x,2), alpha=.3)+
  theme(legend.position = "none")+
  labs(title = "Cooperative + Lang", subtitle="Number of words", y="Number of words", x="Round number", color="playerId")

d.chat %>% 
         left_join(d.games %>% select(gameId, createdAt,condition, chatEnabled)) %>% 
         filter(condition == "competCartel",
                chatEnabled == "true") %>%
ggplot(aes(x = repNum, y=total_num_words, color = as.factor(playerId))) + 
  geom_jitter(alpha = .5) + 
  facet_grid(cols = vars(blockNum), rows = vars(gameId), scales = "free_y") + 
  geom_smooth(method=glm, formula=y~poly(x,2), alpha=.3)+
  theme(legend.position = "none")+
  labs(title="Competitive + Lang", subtitle = "Number of words", y="Number of words", x="Round number", color="playerId")
```

Accuracy
sum utility for entire group/max possible utility per group (or, the max flower *3)
Each condition is going to have a different maximum utility

coopcartel will have a max utility = max flower * 3 (players) * 3 (incentive)

competcartel will have a max utility = sum(top_3_flowers)
```{r}
d.max_utility <- d.contexts %>% 
  left_join(d.games %>% select(gameId, condition)) %>%
  mutate(utility= as.numeric(utility)) %>%
  group_by(gameId, blockNum, repNum) %>% 
  slice_max(order_by =utility,n = 3)%>%
  summarize(competCartel = sum(utility),
            coopCartel = max(utility * 9)) %>% ungroup() %>%
  pivot_longer(cols = c(competCartel, coopCartel), names_to = "condition", values_to = "maxUtility")

d.utility <- d.round_results %>%
         left_join(d.games %>% select(gameId, condition, chatEnabled)) %>% 
  group_by(gameId, blockNum, repNum, trialNum,condition, chatEnabled) %>% 
  summarize(group_utility = sum(as.numeric(playerUtility)))%>% ungroup() %>%
  left_join(d.max_utility) %>%
  mutate(prop_utility = group_utility/maxUtility,
         chat = if_else(chatEnabled=="true", "Lang", "Nonlang"))

d.utility %>% 
  ggplot(aes(x = trialNum, y=prop_utility, color = as.factor(gameId))) + 
  geom_jitter(alpha = .4) + 
  facet_grid(cols = vars(condition), rows = vars(chat), scales = "free_y") + 
  #geom_smooth(method=glm, formula=y~poly(x,2), alpha=.3)+
  theme(legend.position = "none")+
  labs(title="Accuracy", y="Group utility/max utility", x="Round number", color="gameId")
```

-----
